---
title: "Impact Factor"
author: "Marco Mello"
date: "24/09/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::knit_hooks$set(inline = function(x) {
prettyNum(x, big.mark=",")
})
```

Impact factor analysis of the journals classified as "Ecology" (JCR) and "Biodiversity" (CAPES).

[Ecological Synthesis Lab](https://marcomellolab.wordpress.com) (SintECO).

Author: Marco A. R. Mello.

See README for further info.


## Summary

1. [Ecology](#ecology)

2. [Biodiversity](#biodiversity)


```{r, echo=FALSE, v}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

```

```{r, echo=FALSE, include=FALSE}
rm(list= ls())
```


```{r, echo=FALSE, include=FALSE}
library("ggplot2")
library("plyr")
library("RColorBrewer")
library("reshape2")
library("tidyr")
library("vegan")
```




## 1. Ecology{#ecology}

Let's examine impact factor data from the journals classified in the category "Ecology" on Journal Citation Reports (JCR).

```{r, echo=FALSE, include=FALSE}
jcreco = read.csv("data/jcr ecology 2018.csv", na = "Not Available",
                  skip = 1, header = T)

neco<-dim(jcreco)[1]
jcreco<-jcreco[1:(neco-2),]
```


```{r, echo=FALSE, include=FALSE}
str(jcreco)
```


```{r, echo=FALSE, include=FALSE}
jcreco.med <- summarise(jcreco, 
                        IFmedian = median(Journal.Impact.Factor, na.rm = T))
eco.med = round(jcreco.med[1,1], digits = 2)

jcreco.qua1 <- summarise(jcreco, 
                         IFqua1 = quantile(Journal.Impact.Factor, na.rm = T))
eco.qua1 = round(jcreco.qua1[2,1], digits = 2)

jcreco.qua2 <- summarise(jcreco, 
                         IFqua2 = quantile(Journal.Impact.Factor, na.rm = T))
eco.qua2 = round(jcreco.qua2[4,1], digits = 2)

jcreco.min <- summarise(jcreco, 
                        IFmin = min(Journal.Impact.Factor, na.rm = T))
eco.min = round(jcreco.min[1,1], digits = 2)

jcreco.max <- summarise(jcreco, 
                        IFmax = max(Journal.Impact.Factor, na.rm = T))
eco.max = round(jcreco.max[1,1], digits = 2)

```

The median impact factor of Ecology journals is `r format(eco.med, scientific=FALSE)`. Most journals vary between the `r format(eco.qua1, scientific=FALSE)` and `r format(eco.qua2, scientific=FALSE)`. The minimum impact factor recorded is `r format(eco.min, scientific=FALSE)` and the maximum is `r format(eco.max, scientific=FALSE)`.


This is the distribution of impact factors of Ecology:

```{r, echo=FALSE, fig.height=7, fig.width=10, cache=TRUE, out.width='\\textwidth', fig.align='center'}
p1 = ggplot(jcreco, aes(x=Journal.Impact.Factor)) +
  geom_histogram(binwidth=.5, alpha=.3, fill = "blue", position="identity") +
  xlab("Impact factor 2018: JCR's Ecology") + 
  ylab("Frequency") +
  theme(axis.text.x = element_text(colour="grey20",size=20,
                                   angle=0,hjust=.5,vjust=.5,face="plain"),
        axis.text.y = element_text(colour="grey20",size=20,
                                   angle=0,hjust=1,vjust=0,face="plain"),
        axis.title.x = element_text(colour="grey20",size=25,
                                    angle=0,hjust=.5,vjust=0,face="plain"),
        axis.title.y = element_text(colour="grey20",size=25,
                                    angle=90,hjust=.5,vjust=.5,face="plain")) +
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5),"cm")) +
  theme(axis.title.y=element_text(margin=margin(0,20,0,0)),
        axis.title.x=element_text(margin=margin(20,0,0,0))) +
  geom_vline(data=jcreco.med, aes(xintercept=eco.med),
             linetype="dashed", size=0.5, color = "blue") +
  geom_text(aes(eco.med, 26, label = eco.med, angle = 0, 
                hjust = -0.1, vjust = 0), size = 5, color = "blue") +
  geom_vline(data=jcreco.qua1, aes(xintercept=eco.qua1),
             linetype="dashed", size=0.5, color = "blue") +
  geom_text(aes(eco.qua1, 26, label = eco.qua1, angle = 0, 
                hjust = -0.1, vjust = 0), size = 3, color = "blue") +
  geom_vline(data=jcreco.qua1, aes(xintercept=eco.qua2),
             linetype="dashed", size=0.5, color = "blue") +
  geom_text(aes(eco.qua2, 26, label = eco.qua2, angle = 0,
                hjust = -0.1, vjust = 0), size = 3, color = "blue")

p1
```


## 2. Biodiversity{#biodiversity}

```{r, echo=FALSE, include=FALSE}
str(jcreco)
```

The journals classified in the categories "Oceanography", "Plant Sciences", and "Zoology" on Journal Citation Reports (JCR). Together with "Ecology", they compose the category "Biodiversity" from the Brazilian Coordination for the Improvement of Higher Education Personnel (CAPES).

```{r, echo=FALSE, include=FALSE}
jcroce = read.csv("data/jcr oceanography 2018.csv", na = "Not Available",
                  skip = 1, header = T)
noce<-dim(jcroce)[1]
jcroce<-jcroce[1:(noce-2),]

jcrpla = read.csv("data/jcr plant sciences 2018.csv", na = "Not Available",
                  skip = 1, header = T)
npla<-dim(jcrpla)[1]
jcrpla<-jcrpla[1:(npla-2),]

jcrzoo = read.csv("data/jcr zoology 2018.csv", na = "Not Available",
                  skip = 1, header = T)
nzoo<-dim(jcrzoo)[1]
jcrzoo<-jcrzoo[1:(nzoo-2),]

```


```{r, echo=FALSE, include=FALSE}
jcrbio <- rbind(jcreco, jcroce[1:nrow(jcroce),], 
                jcrpla[1:nrow(jcrpla),], jcrzoo[1:nrow(jcrzoo),])
```


```{r, echo=FALSE, include=FALSE}
str(jcrbio)
```


```{r, echo=FALSE, include=FALSE}
jcrbio$Category <- c(replicate(nrow(jcreco), "Ecology"), 
                     replicate(nrow(jcroce), "Oceanography"), 
                     replicate(nrow(jcrpla), "PlantScience"), 
                     replicate(nrow(jcrzoo), "Zoology"))
```


```{r, echo=FALSE, include=FALSE}
jcrbio.med <- ddply(jcrbio, "Category", summarise, 
                    IFmedian = median(Journal.Impact.Factor, na.rm = T))
bio.med <- summarise(jcrbio, 
                        IFmedian = median(Journal.Impact.Factor, na.rm = T))
bio.med = round(bio.med, digits = 2)

jcrbio.qua1 <- summarise(jcrbio, 
                         IFqua1 = quantile(Journal.Impact.Factor, na.rm = T))
bio.qua1 = round(jcrbio.qua1[2,1], digits = 2)

jcrbio.qua2 <- summarise(jcrbio, 
                         IFqua2 = quantile(Journal.Impact.Factor, na.rm = T))
bio.qua2 = round(jcrbio.qua2[4,1], digits = 2)

jcrbio.min <- summarise(jcrbio, 
                        IFmin = min(Journal.Impact.Factor, na.rm = T))
bio.min = round(jcrbio.min[1,1], digits = 2)

jcrbio.max <- summarise(jcrbio, 
                        IFmax = max(Journal.Impact.Factor, na.rm = T))
bio.max = round(jcrbio.max[1,1], digits = 2)
```


The median impact factor of Biodiversity journals is `r format(bio.med, scientific=FALSE)`. Most journals vary between the `r format(bio.qua1, scientific=FALSE)` and `r format(bio.qua2, scientific=FALSE)`. The minimum impact factor rjcrbiorded is `r format(bio.min, scientific=FALSE)` and the maximum is `r format(bio.max, scientific=FALSE)`.


```{r, echo=FALSE, include=FALSE}
eco.med = round(jcrbio.med[1,2], digits = 2)
oce.med = round(jcrbio.med[2,2], digits = 2)
pla.med = round(jcrbio.med[3,2], digits = 2)
zoo.med = round(jcrbio.med[4,2], digits = 2)
```


These are the median impact factors by category:

1. Ecology = `r format(eco.med, scientific=FALSE)`

2. Oceanography = `r format(oce.med, scientific=FALSE)`

3. Plant Sciences = `r format(pla.med, scientific=FALSE)`

4. Zoology = `r format(zoo.med, scientific=FALSE)`


This is the distribution of impact factors by category:

```{r, echo=FALSE, warning=FALSE, fig.height=7, fig.width=10, cache=TRUE, out.width='\\textwidth', fig.align='center'}
p2 = ggplot(jcrbio, aes(x=Journal.Impact.Factor, fill=Category)) +
  geom_density(alpha=.3, position="identity", color = F) +
  xlab("Impact factor 2018: CAPES's Biodiversity") + 
  ylab("Frequency") +
  theme(axis.text.x = element_text(colour="grey20",size=20,angle=0,
                                   hjust=.5,vjust=.5,face="plain"),
        axis.text.y = element_text(colour="grey20",size=20,angle=0,
                                   hjust=1,vjust=0,face="plain"),
        axis.title.x = element_text(colour="grey20",size=25,angle=0,
                                    hjust=.5,vjust=0,face="plain"),
        axis.title.y = element_text(colour="grey20",size=25,angle=90,
                                    hjust=.5,vjust=.5,face="plain")) +
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5),"cm")) +
  theme(axis.title.y=element_text(margin=margin(0,20,0,0)),
        axis.title.x=element_text(margin=margin(20,0,0,0))) +
  geom_vline(data=jcrbio.med, aes(xintercept=IFmedian,  colour=Category),
             linetype="dashed", size=0.5) +
  geom_text(aes(eco.med, 0.60, label = eco.med, angle = 0, 
                hjust = 0, vjust = 0), size = 3) +
  geom_text(aes(oce.med, 0.62, label = oce.med, angle = 0, 
                hjust = 0, vjust = 0), size = 3) +
  geom_text(aes(pla.med, 0.64, label = pla.med, angle = 0, 
                hjust = 0, vjust = 0), size = 3) +
  geom_text(aes(zoo.med, 0.66, label = zoo.med, angle = 0, 
                hjust = 0, vjust = 0), size = 3)

p2
```

